---
title: "Logistic Regression Homework"
output: html_document
date: "2024-08-20"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(car)
library(stats)
library(UsingR)
library(AmesHousing)
library(car)
library(DescTools)
library(corrplot)
library(mgcv)
library(visreg)
library(DataExplorer)
```

```{r}
train <- read.csv('insurance_t.csv')
test <- read.csv('insurance_v.csv')
```

**Phase 1**
Binary Variables:
```{r}
#Select binary variables
binary_vars <- c("DDA", "DIRDEP", "NSF", "SAV", "ATM", "CD", "IRA", "LOC", "INV", "ILS", "MM", "MTG", "CC", "SDB", "HMOWN", "MOVED", "INAREA", "INS")

# Apply Mantel-Haenszel Test
binary_results <- sapply(binary_vars, function(var) {
  test <- CMHtest(table(train[[var]], train$INS))$table[1,]
  return(c(var,  test[1], test[3]))
})

# Convert to data frame and rank by p-value
binary_results_df <- as.data.frame(t(binary_results))
names(binary_results_df) <- c("Variable", "Chi-Square Statistic", "P-Value")
binary_results_df <- binary_results_df %>% arrange(as.numeric(`P-Value`))
binary_results_df[-1,]
```


Ordinal Variables:
```{r}
#Select ordinal variables
ordinal_vars <- c("CASHBK", "MMCRED", "CCPURC")  

# Apply Mantel-Haenszel Test
ordinal_results <- sapply(ordinal_vars, function(var) {
  test <- CMHtest(table(train[[var]], train$INS))$table[1,]
  return(c(var,  test[1], test[3]))
})

# Convert to data frame and rank by p-value
ordinal_results_df <- as.data.frame(t(ordinal_results))
names(ordinal_results_df) <- c("Variable", "Correlation", "P-Value")
ordinal_results_df <- ordinal_results_df %>% arrange(as.numeric(`P-Value`))
ordinal_results_df
```


Nominal Variables:
```{r}
#Select Nominal Variables
nominal_vars <- c("BRANCH", "RES")

#Apply Chi Sq Test
nominal_results <- sapply(nominal_vars, function(var) {
  test <- chisq.test(table(train[[var]], train$INS))
  return(c(var, test$statistic, test$p.value))
})

# Convert to data frame and rank by p-value
nominal_results_df <- as.data.frame(t(nominal_results))
names(nominal_results_df) <- c("Variable", "Chi-Square Statistic", "P-Value")
nominal_results_df <- nominal_results_df %>% arrange(as.numeric(`P-Value`))
nominal_results_df
```

Continuous Variables:
```{r}
#Determine the variables with more than 10 distinct values
names(train[sapply(train, function(x) n_distinct(x))>10])

#Select continuous variables
continuous_vars <- c("ACCTAGE", "DDABAL", "DEP", "DEPAMT", "CHECKS", "NSFAMT", "PHONE", "TELLER", "SAVBAL", "ATMAMT", "POS", "POSAMT", "CDBAL", "IRABAL", "LOCBAL", "INVBAL", "ILSBAL", "MMBAL", "MTGBAL", "CCBAL", "INCOME", "LORES", "HMVAL", "AGE" , "CRSCORE")

# Apply t-test
continuous_results <- sapply(continuous_vars, function(var) {
  formula <- as.formula(paste("INS ~", var))
  mod <- glm(formula, data=train, family=binomial(link="logit"))
  test <- unname(summary(mod)$coefficients[,4])[2]
  return(c(var, test))
})

# Convert to data frame and rank by p-value
continuous_results_df <- as.data.frame(t(continuous_results))
names(continuous_results_df) <- c("Variable", "P-Value")
continuous_results_df <- continuous_results_df %>% arrange(as.numeric(`P-Value`))
continuous_results_df
```

```{r}
# Combine results into one table
all_results_df <- bind_rows(
  binary_results_df %>% mutate(Type = "Binary"),
  ordinal_results_df %>% mutate(Type = "Ordinal"),
  nominal_results_df %>% mutate(Type = "Nominal"),
  continuous_results_df %>% mutate(Type = "Continuous")
) %>% dplyr::select('Variable', 'P-Value', 'Type')

# Arrange by P-Value (ascending)
all_results_df <- all_results_df %>% arrange(Type, as.numeric(`P-Value`))

# View the summarized table that includes all variables
print(all_results_df[-1,])

#Only include significant variables
signif = all_results_df %>% subset(as.numeric(`P-Value`) < 0.002)
print(signif[-1,])
```

Odds Ratios:
```{r}
# Find Odds Ratios
Odds_Ratios <- sapply(binary_vars, function(var) {
  test <- OddsRatio(table(train[[var]], train$INS))
  return(c(var, test))
})

# Convert to data frame and rank by p-value
odds_ratio_df <- as.data.frame(t(Odds_Ratios))
names(odds_ratio_df) <- c("Variable", "Odds Ratio")
odds_ratio_df$`Odds Ratio` = as.numeric(odds_ratio_df$`Odds Ratio`)
odds_ratio_df <- odds_ratio_df %>% arrange(desc(`Odds Ratio`))
odds_ratio_df[-1,]
```
The odds of a customer purchasing the insurance product are 3.47 times higher than not purchasing the insurance product when the customer has an investment account.

Linearity Assumption:
```{r}
#Create GAM models
for (var in continuous_vars){
formula <- as.formula(paste("INS ~ s(", var, ")"))
fit.gam <- mgcv::gam(formula, data=train, family=binomial(link="logit"), method = 'REML')
print(summary(fit.gam))
}
```

```{r}
#Plot GAM models
for (var in continuous_vars){
train = train %>% mutate(CCPURC = ifelse(is.na(CCPURC), 0, CCPURC))
formula <- as.formula(paste("INS ~ s(", var, ")"))
fit.gam <- mgcv::gam(formula, data=train, family=binomial(link="logit"), method = 'REML')
plot(fit.gam, all.terms = TRUE)
}
```

```{r}
#See if spline adds value
for (var in continuous_vars){
formula <- as.formula(paste("INS ~ s(", var, ")"))
fit.gam <- mgcv::gam(formula, data=train, family=binomial(link="logit"), method = 'REML')
formula2 <- as.formula(paste("INS ~", var))
logit.model <- glm(formula2, data=train, family=binomial(link="logit"))
print(anova(logit.model, fit.gam, test="LRT"))
}
```

Linear: ACCTAGE, CASHBK, LOCBAL, MTGBAL, CCBAL, INCOME, LORES, HMVAL, AGE, CRSCORE



Missing Values:
```{r}
#Determine which variables have the highest amount of missing values
sort(colSums(is.na(train)), decreasing = TRUE)
```
```{r}
missing_data <- train %>% dplyr::select('AGE', 'INCOME', 'LORES', 'HMVAL', 'HMOWN', 'PHONE', 'POS', 'POSAMT', 'INV', 'INVBAL', 'CC', 'CCBAL', 'CC') %>%
  summarise_all(~sum(is.na(.))) %>%
  gather(key = "Variable", value = "MissingValues")

# Arrange the variables in descending order of missing values
missing_data <- missing_data %>%
  arrange(desc(MissingValues))

# Create a bar plot to visualize missing values
ggplot(missing_data, aes(x = reorder(Variable, -MissingValues), y = MissingValues)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Missing Values per Variable", x = "Variables", y = "Number of Missing Values") +
  coord_flip() +
  theme_minimal()
```


Redundant Variables:
DDA and DDABAL
SAV and SAVBAL
ATM and ATMAMT
CD and CDBAL
IRA and IRABAL
LOC and LOCBAL
INV and INVBAL
ILS and ILSBAL
MM and MMBAL
MTG and MTGBAL
CC and CCBAL
POS and POSAMT
DDABAL and DEPAMT
HV and HMOWN





